---
title: "Working with skills"
vignette: >
  %\VignetteIndexEntry{Working with skills}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

## What is a skill?

A skill is a folder with a `SKILL.md` file inside it. The file starts with a short header giving the skill a name and a one-sentence description of when it applies, followed by instructions written in plain Markdown. Some skills also carry supporting files alongside `SKILL.md`: reference documents, templates, or worked examples the instructions point to.

AI coding assistants such as Claude Code and Codex look for skills in a known folder in your project. When you ask for something that matches a skill's description, the assistant reads the instructions and follows them. Without a skill, the assistant has to guess how you like things done. With one, it knows.

Posit's [Quarto authoring skill](https://github.com/posit-dev/skills/blob/main/quarto/quarto-authoring/SKILL.md) is a good example. Its description says to use it whenever you are working with `.qmd` files or Quarto projects, and its instructions and reference files cover callouts, cross-references, citations, and the other details an assistant tends to get wrong.

Skills live on GitHub, but installing one usually means `npm` or copying folders by hand. {ally} does it from R.

## Installation

```{r}
#| eval: false
# install.packages("pak")
pak::pak("rfortherestofus/ally")
```

## Install your first skill

Open the project you want the skill in, copy the link to the skill's `SKILL.md` from your browser, and paste it into `install_skill()`:

```{r}
#| eval: false
library(ally)

install_skill(
  "https://github.com/posit-dev/skills/blob/main/quarto/quarto-authoring/SKILL.md"
)
#> ✔ Installed "quarto-authoring" to '~/projects/report/.agents/skills/quarto-authoring'
#> ℹ Codex and other agents read '.agents/skills/' directly.
#> ✔ Copied to "Claude Code" ('~/projects/report/.claude/skills')
```

That is the whole workflow. Open the project in Claude Code or Codex, ask for help with a Quarto document, and the assistant picks the skill up on its own.

## Where the files go

`install_skill()` puts the skill in two places inside the project:

| Path | What it is |
|---|---|
| `.agents/skills/<skill>/` | The canonical copy. Codex and a growing number of agents read it directly. |
| `.claude/skills/<skill>/` | A second copy for Claude Code, which only looks in its own folder. |

Both folders are plain directories, so they can be committed to git and shared with collaborators.

## A walkthrough you can run

Everything below runs against a throwaway project and a skill written on the fly, so it works offline and touches nothing else on your computer.

```{r}
library(ally)

# A minimal skill: a folder with a SKILL.md inside it
skill <- fs::dir_create(fs::path(tempdir(), "r-style-guide"))
writeLines(
  c(
    "---",
    "name: r-style-guide",
    "description: House style for R code in this project.",
    "---",
    "",
    "Follow the tidyverse style guide. Use the base pipe `|>`."
  ),
  fs::path(skill, "SKILL.md")
)

# An empty project to install it into
project <- fs::dir_create(fs::path(tempdir(), "my-project"))
```

Install from the local path. The project is set as the working directory just for this call:

```{r}
withr::with_dir(project, install_skill(skill))
```

Both copies are now in place:

```{r}
fs::dir_tree(project, all = TRUE)
```

`installed_skills()` lists what is installed in a project:

```{r}
withr::with_dir(project, installed_skills())
```

And `remove_skill()` deletes both copies:

```{r}
withr::with_dir(project, remove_skill("r-style-guide"))
```

## This project or every project

By default a skill is installed into the current working directory, so it travels with the project. Pass `scope = "user"` to install it into your home folder instead, where every project on the computer can use it:

```{r}
#| eval: false
install_skill("posit-dev/skills/quarto/quarto-authoring", scope = "user")
#> ✔ Installed "quarto-authoring" to '~/.agents/skills/quarto-authoring'
#> ℹ Every project on this computer can use skills in '~/.agents/skills/'.
#> ✔ Copied to "Claude Code" ('~/.claude/skills')
```

Every function in {ally} takes the same `scope` argument.

## Ways to name a skill

`install_skill()` accepts a few forms of `source`:

```{r}
#| eval: false
# A full GitHub URL, pasted straight from the browser
install_skill(
  "https://github.com/posit-dev/skills/tree/main/quarto/quarto-authoring"
)

# The link to the SKILL.md itself also works
install_skill(
  "https://github.com/posit-dev/skills/blob/main/quarto/quarto-authoring/SKILL.md"
)

# A repository whose SKILL.md sits at the top level is named after the repository
install_skill("https://github.com/statzhero/tidy-r-skill")

# GitHub shorthand: owner/repo/path/to/skill, with an optional @ref
install_skill("posit-dev/skills/quarto/quarto-authoring")
install_skill("owner/repo/skills/my-skill@dev")

# A local path, handy while writing a skill of your own
install_skill("~/my-skills/r-style-guide")
```

GitHub skills are fetched from the repository's zip archive: a single download that needs no token and never counts against the GitHub API rate limit, so a room full of people on one network can install at once. Public repositories only.

## Keeping skills current

Skills change. Authors fix mistakes, add reference material, and adapt their instructions as the tools they describe evolve. The Quarto authoring skill, for instance, records a version number in its `SKILL.md` header and has been revised several times. An installed skill is a snapshot of the moment you installed it; it does not update itself.

{ally} records where each skill came from in a small `.ally-source.json` file next to the `SKILL.md`, so getting the newest version is one call:

```{r}
#| eval: false
update_skill("quarto-authoring")
#> ℹ Updating "quarto-authoring" from "posit-dev/skills/quarto/quarto-authoring@main"
#> ✔ Installed "quarto-authoring" to '~/projects/report/.agents/skills/quarto-authoring'
#> ℹ Codex and other agents read '.agents/skills/' directly.
#> ✔ Copied to "Claude Code" ('~/projects/report/.claude/skills')
```

`update_skill()` fetches the skill again from its original source, replaces the canonical copy, and refreshes the Claude Code copy. To bring every skill in a project up to date at once:

```{r}
#| eval: false
for (skill in installed_skills()) {
  update_skill(skill)
}
```

Skills installed with `scope = "user"` are updated the same way, with `scope = "user"` added to each call.

If you edit a skill by hand in `.agents/skills/`, or delete the Claude Code copy by accident, `link_skills()` refreshes the copies for every installed skill without fetching anything:

```{r}
#| eval: false
link_skills()
```

Prefer a single set of files? Pass `link = TRUE` to `install_skill()`, `update_skill()`, or `link_skills()` and {ally} puts a relative symlink at `.claude/skills/<skill>` pointing back to the canonical copy instead of a second copy. Where symlinks are not supported (typically Windows without developer mode) it falls back to copying and tells you so.

## Where to find skills

- [Ally](https://rfortherestofus.com/ally) from R for the Rest of Us offers a curated collection of skills for R users working with AI. Each one comes with the link to paste into `install_skill()`.
- [posit-dev/skills](https://github.com/posit-dev/skills) is Posit's collection, covering R packages, Shiny, Quarto, and more.
- Write your own. A folder with a `SKILL.md` is all it takes, and `install_skill()` accepts a local path while you iterate.
